---
title: "STA 380 HW2"
author: "Jack Cheng"
date: "8/15/2017"
output:
  rmarkdown: github_document
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```


# Question 1 Flights at ABIA

### Goal: Our team decided to answer the following question: when is the best time in a year to fly from Austin to other cities?

### Data cleaning

```{r}
library(ggplot2)
ABIA = read.csv("~/Documents/STA380/ABIA.csv")
```


As a first step, we did a four-step data cleaning： dropped all the missing values in *ArrDelay* and *DepDelay*, converted all months and day of week numbers into factors with names, selected only filghts originated from Austin,  and lastly summed the departure delay time and the arrival delay time together to become the total delay time. 

```{r}
# We dropped missing values and converted months and days into factors with month and day of week names. We only kept flights originated from Austin.

ABIA_cleaned = ABIA[!is.na(ABIA$ArrDelay),][!is.na(ABIA$DepDelay),][ABIA$Origin=='AUS',]
ABIA_cleaned$Month<-factor(ABIA_cleaned$Month,levels=c(1,2,3,4,5,6,7,8,9,10,11,12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
ABIA_cleaned$DayOfWeek<-factor(ABIA_cleaned$DayOfWeek,levels=c(1,2,3,4,5,6,7),labels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"))

# For these remaining flights, we summed the departure and arrival delay times together to get the total delay time. We believe the total delay time is the one which travelers care the most.

ABIA_cleaned$TotDelay=ABIA_cleaned$ArrDelay+ABIA_cleaned$DepDelay

```

### Analysis by month

We computed and plotted the average total delay time in each month

```{r}
month_delay=ABIA_cleaned[,c('Month','TotDelay','DayOfWeek')]
month_avg_delay = aggregate(.~Month, data=month_delay, mean, na.rm=TRUE, na.action=NULL)
ggplot(data = month_avg_delay,mapping = aes(x= Month, y= TotDelay )) + geom_col()

```

As can be seen from the graphs above, March, June, and December are three months with highest average total delay time. We would recommend you to avoid travelling in these months if possible. 

### Analysis by the day of week in each month

However, we understand that it is often inconvenient to make such dramatic change to your travel plan. Our team wanted to know if it is possible to change the day of the week to travel in those months with highest average total delay time. So we did a further analysis on average delay time in those three months based on different days of a week.

```{r}
new = aggregate(month_delay$TotDelay, by=list(Month=month_delay$Month, DayOfWeek=month_delay$DayOfWeek), FUN=mean)

list= c("Mar","Jun","Dec")

par(mfrow=c(1,3))
for (i in 1:length(list)){
  mon=new[which(new[,1]==list[i]),]
  monthstr=list[i]
  xlab=paste('Day of the Week in',monthstr)
  print(ggplot(data = mon,mapping = aes(x= DayOfWeek, y= x))+geom_col()+labs(x=xlab, y='Average Delay Time'))
  
}


```

As can be seen from the plots above, a good choice is to travel on Wednesday since it has the lowest average delay time in a week. 


### Analysis by distance 

My team also wanted to test if there is a difference between short and long distance flights across all 12 months.It is commonly known that it is easier for long distance (more than 750 miles) flights to re-accommodate their time by changing their speed. This kind of re-accommodating may require certain traffic and weather conditions that are related to month.

```{r}
LongFlight=ABIA_cleaned[which(ABIA_cleaned$Distance>=750),c('Month','TotDelay','DayOfWeek')]
ShortFlight=ABIA_cleaned[which(ABIA_cleaned$Distance<750),c('Month','TotDelay','DayOfWeek')]



LongFlight_month = aggregate(.~Month, data=LongFlight, mean, na.rm=TRUE, na.action=NULL)
ShortFlight_month = aggregate(.~Month, data=ShortFlight, mean, na.rm=TRUE, na.action=NULL)

LongFlight_month$LongFlightSave=LongFlight_month$TotDelay-ShortFlight_month$TotDelay

LongFlight_month[which(LongFlight_month$LongFlightSave>=0),c(1,4)]
```

So for the months listed above, we recommend taking a long distance trip in order to minimize the delay. As for the rest of the year, a short distance trip would be better.

### Conclusion 

Given the analysis above, our team offer these suggestions to the travelers in Austin:

* If possible, schedule your trip in September, October and December to avoid the heavy traffic and long delay time.

* If you have to travel among those months, Wednesday would be recommended for the sake of attaining relatively lower delay time.

* One additional suggestion: March, May, and October are the ideal months for short trips, and also, the rest months of the year are ideal months for longer trips.
