---
title: "STA 380 HW2"
author: "Jack Cheng"
date: "8/15/2017"
output:
  rmarkdown: github_document
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

```{r}
library(ggplot2)
ABIA = read.csv("~/Documents/STA380/ABIA.csv")
```
Our team decided to answer the following question: when is the best time in a year to fly from Austin to other cities?

As a first step, we did data cleaning and dropped all the missing values in ArrDelay and DepDelay since we only care about the data that has a status of delay. Then, we converted all months and day of week numbers to factors with names. Lastly, we summed the departural delay time and the arrival delay time together to become the total delay time. 

```{r}
# Data cleaning: drop missing values and convert months and days into factors with month and day of week names.
ABIA_cleaned = ABIA[!is.na(ABIA$ArrDelay),][!is.na(ABIA$DepDelay),][ABIA$Origin=='AUS',]
ABIA_cleaned$Month<-factor(ABIA_cleaned$Month,levels=c(1,2,3,4,5,6,7,8,9,10,11,12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
ABIA_cleaned$DayOfWeek<-factor(ABIA_cleaned$DayOfWeek,levels=c(1,2,3,4,5,6,7),labels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"))

# For these delayed flights, we sum up the departure and arrival delay to get the total delay. We believe that this is the delay people care about the most.
ABIA_cleaned$TotDelay=ABIA_cleaned$ArrDelay+ABIA_cleaned$DepDelay
month_delay=ABIA_cleaned[,c('Month','TotDelay','DayOfWeek')]
month_avg_delay = aggregate(.~Month, data=month_delay, mean, na.rm=TRUE, na.action=NULL)
ggplot(data = month_avg_delay,mapping = aes(x= Month, y= TotDelay )) + geom_col()

```

As we can see from the graph of the departure and arrival delay,March, June, and December are three months with really high average total delays. The reason behind this may be that many students in Austin seize this chance to go back home or travel during spring break, summer holiday and Christmas times.

For the reasons above, we strongly suggest you to arrange your travel in months such as September, October, and November to minimize the delay.

However, in the real case, we are limited with out choice in terms of flying dates and time choosing. Our team's advice is that it is a great idea for UT students to take into consideration of these delay issues and arrange a better time to travel.Thus, we did a further analysis on the average delays in these three months based on days in a week.

```{r}
new = aggregate(month_delay$TotDelay, by=list(Month=month_delay$Month, DayOfWeek=month_delay$DayOfWeek), FUN=mean)

```

```{r}
list= c("Mar","Jun","Dec")

par(mfrow=c(1,3))
for (i in 1:length(list)){
  mon=new[which(new[,1]==list[i]),]
  monthstr=list[i]
  xlab=paste('Day of the Week in',monthstr)
  print(ggplot(data = mon,mapping = aes(x= DayOfWeek, y= x))+geom_col()+labs(x=xlab, y='Average Delay Time'))
  
}


```

A good choice here is to arrange the trip on Wednesdays since it isthe day with the lowest average delay in a week. My team also held an assumption that the flights for different distance would be influenced differently in different months. Due to weather and traffic factors, it would be easier for those who travel long distance ( more than 750 miles) to re-accommodate their time by changing their speed at certain months. This kind of re-accommodating requires certain heights and weather conditions related to months.


```{r}
LongFlight=ABIA_cleaned[which(ABIA_cleaned$Distance>=750),c('Month','TotDelay','DayOfWeek')]
ShortFlight=ABIA_cleaned[which(ABIA_cleaned$Distance<750),c('Month','TotDelay','DayOfWeek')]



LongFlight_month = aggregate(.~Month, data=LongFlight, mean, na.rm=TRUE, na.action=NULL)
ShortFlight_month = aggregate(.~Month, data=ShortFlight, mean, na.rm=TRUE, na.action=NULL)

LongFlight_month$LongFlightSave=LongFlight_month$TotDelay-ShortFlight_month$TotDelay

LongFlight_month[which(LongFlight_month$LongFlightSave>=0),c(1,4)]
```

So for the months listed above, we recommend that it is suggestive to take a long distance trip in order to mnimize the delay. As for the rest of the year, a short distance trip would be better.


